Preferences for the commodification of pensions in Chile: the rol of intergenerational social mobility

### 2025-11-03 paper/04-analysis.qmd
Update entrega 1

---
output: html_document
editor_options: 
  chunk_output_type: console
---

@@

```{r}
#| label: libraries2
#| message: false
#| echo: false

if (!require("pacman")) install.packages("pacman")

pacman::p_load(tidyverse,
               here,
               sjPlot,
               sjmisc,
               texreg,
               estimatr,
               conflicted,
               ggdist,
               summarytools,
               dagitty,
               ggdag)

conflicts_prefer(dplyr::filter)

options(scipen=999)
rm(list = ls())
```

```{r}
#| label: data2
#| message: false
#| echo: false

load(here("input/data/proc/data_v2.3.RData"))

@@

```{r}
#| label: fig-alluvial
#| fig-cap: "Change in preferences for pension commodification over time (2016, 2018, and 2023)."
#| fig-cap-location: bottom
#| echo: false
#| results: asis
db %>% 
  mutate(just_pension_f = factor(just_pension_f, levels = c(0,1), labels = c("Disagree", "Agree"))) %>%
  group_by(wave, just_pension_f) %>% 
  count() %>% 
  group_by(wave) %>% 
  mutate(porcentaje=n/sum(n)) %>% 
  ungroup() %>% 
  ggplot(aes(x = wave, y = porcentaje, group = just_pension_f)) +
  geom_col(aes(fill = just_pension_f)) +
  scale_y_continuous(labels = scales::percent) + 
  scale_fill_manual(
    values = c("#5b59d4", "#c8237d"),
    breaks = c("Agree","Disagree")
  )  +
  coord_flip() +
  geom_text(
    aes(label = paste0(round(porcentaje * 100, 1), "%")),
    position = position_stack(vjust = 0.5),
    color = "white",
    size = 3,
    fontface = "bold"
  ) +
  labs(y = NULL,
       x = NULL,
       fill = NULL,
       caption = "Source: own elaboration with data from ELSOC 2016-2023 (N obs = 3,435)") +
  theme_ggdist() +
  theme(legend.position = "top") 
```

@@

pdat <- db %>% 
  select(wave, estrato_orig, estrato_ocupa, just_pension_f) %>% 
  drop_na() %>% 
  mutate(
    mobility = paste(estrato_orig, estrato_ocupa, sep = "-"),
    facet = case_when(
      mobility %in% c("High-Middle","High-Low","Middle-Low") ~ "Downward",
      estrato_orig == estrato_ocupa ~ "Inmobile",
      mobility %in% c("Middle-High","Low-Middle","Low-High") ~ "Upward"
    )
  ) %>% 
  group_by(wave, mobility, facet) %>% 
  summarise(y = mean(just_pension_f), .groups = "drop") %>% 
  group_by(mobility, facet) %>% 
  mutate(gm = mean(y)) %>% 
  ungroup()

# datos SOLO para el punto (uno por categoría)
pdat_gm <- pdat %>% distinct(mobility, facet, gm)

ggplot(pdat, aes(x = mobility, y = y, group = wave)) +
  geom_col(aes(fill = wave), position = "dodge", alpha = 0.9) +
  # Punto "X" (pch 4) con su propia leyenda (shape)
  geom_point(
    data = pdat_gm,
    aes(x = mobility, y = gm, shape = "All waves"),
    inherit.aes = FALSE,
    size = 3, stroke = 1, color = "black"
  ) +
  scale_y_continuous(labels = scales::percent, limits = c(0, 1)) + 
  scale_fill_manual(
    values = c("2016" = "#5b59d4", "2018" = "#b3b3b3ff", "2023" = "#c8237d"),
    name   = "Wave"
  ) +
  scale_shape_manual(                     # <-- leyenda separada para el shape
    name = NULL,                          # o "Marker" si quieres título
    values = c("All waves" = 4)          # 4 = "X"
  ) +
  guides(
    fill  = guide_legend(order = 1),
    shape = guide_legend(order = 2, override.aes = list(color = "black"))
  ) +
  facet_wrap(~ facet, scales = "free_x") +
  labs(
    y = "% of Agreement",
    x = NULL,
    caption = "Source: own elaboration with data from ELSOC 2016–2023 (N obs = 3,435)"
  ) +
  theme_ggdist() +
  theme(legend.position = "top")

@@

@fig-reg reports the doubly robust estimates of the effect of intergenerational mobility on preferences for pension commodification, obtained from weighted models combining stabilized IPW–ATT with covariate adjustment and wave fixed effects. This specification ensures consistency if either the weighting or the outcome model is correctly specified, while improving efficiency [@wooldridge_introductory_2009]. Results are virtually identical to baseline models without covariates (see [Supplementary Material](#anexo) for complete models), with no substantive change in the size or significance of the mobility coefficients.

```{r}
#| label: models
#| echo: false
#| results: false

### Pensiones

### Low to middle

data_lm <- datos1 %>% 
  select(idencuesta, 
         ola, 
         y = just_pension_f, 
         t = moblm, 
         w = w_low_lm_t, 
         edad = m0_edad_f, 
         sexo = sexo_f, 
         nac = nacionalidad_f,
         etnia = etnia_f,
         hogar = hogar_bip_f, 
         edu0 = educ_orig_f) %>% 
  filter(!is.na(t), !is.na(w)) %>% 
  mutate(ola = factor(ola, levels = c("1", "2", "3"),
                      labels = c("2016", "2018", "2023")))

mp_lm <-
  lm_robust(y ~ t + edad + sexo + nac + etnia + hogar + edu0 + ola,
            weights = w,
            se_type = "CR2",
            clusters = idencuesta,
            data = data_lm)

### Low to high

data_lh <- datos1 %>% 
  select(idencuesta, 
         ola, 
         y = just_pension_f, 
         t = moblh, 
         w = w_low_lh_t, 
         edad = m0_edad_f, 
         sexo = sexo_f, 
         nac = nacionalidad_f,
         etnia = etnia_f,
         hogar = hogar_bip_f, 
         edu0 = educ_orig_f) %>% 
  filter(!is.na(t), !is.na(w)) %>% 
  mutate(ola = factor(ola, levels = c("1", "2", "3"),
                      labels = c("2016", "2018", "2023")))

mp_lh <- 
  lm_robust(y ~ t + edad + sexo + nac + etnia + hogar + edu0+ ola,
            weights = w,
            se_type = "CR2",
            clusters = idencuesta,
            data = data_lh)

### middle to high

data_mh <- datos2 %>% 
  select(idencuesta, 
         ola, 
         y = just_pension_f, 
         t = mobmh, 
         w = w_mid_mh_t, 
         edad = m0_edad_f, 
         sexo = sexo_f, 
         nac = nacionalidad_f,
         etnia = etnia_f,
         hogar = hogar_bip_f, 
         edu0 = educ_orig_f) %>% 
  filter(!is.na(t), !is.na(w)) %>% 
  mutate(ola = factor(ola, levels = c("1", "2", "3"),
                      labels = c("2016", "2018", "2023")))

mp_mh <-  
  lm_robust(y ~ t + edad + sexo + nac + etnia + hogar + edu0+ ola,
            weights = w,
            se_type = "CR2",
            clusters = idencuesta,
            data = data_mh)

### middle to low

data_ml <- datos2 %>% 
  select(idencuesta, 
         ola, 
         y = just_pension_f, 
         t = mobml, 
         w = w_mid_ml_t, 
         edad = m0_edad_f, 
         sexo = sexo_f, 
         nac = nacionalidad_f,
         etnia = etnia_f,
         hogar = hogar_bip_f, 
         edu0 = educ_orig_f) %>% 
  filter(!is.na(t), !is.na(w)) %>% 
  mutate(ola = factor(ola, levels = c("1", "2", "3"),
                      labels = c("2016", "2018", "2023")))

mp_ml <- 
  lm_robust(y ~ t + edad + sexo + nac + etnia + hogar + edu0+ ola,
            weights = w,
            se_type = "CR2",
            clusters = idencuesta,
            data = data_ml)

### High to middle

data_hm <- datos3 %>% 
  select(idencuesta, 
         ola, 
         y = just_pension_f, 
         t = mobhm, 
         w = w_hig_hm_t, 
         edad = m0_edad_f, 
         sexo = sexo_f, 
         nac = nacionalidad_f,
         etnia = etnia_f,
         hogar = hogar_bip_f, 
         edu0 = educ_orig_f) %>% 
  filter(!is.na(t), !is.na(w)) %>% 
  mutate(ola = factor(ola, levels = c("1", "2", "3"),
                      labels = c("2016", "2018", "2023")))

mp_hm <- 
  lm_robust(y ~ t + edad + sexo + nac + etnia + hogar + edu0+ ola,
            weights = w,
            se_type = "CR2",
            clusters = idencuesta,
            data = data_hm)

### High to low

data_hl <- datos3 %>% 
  select(idencuesta, 
         ola, 
         y = just_pension_f, 
         t = mobhl, 
         w = w_hig_hl_t, 
         edad = m0_edad_f, 
         sexo = sexo_f, 
         nac = nacionalidad_f,
         etnia = etnia_f,
         hogar = hogar_bip_f, 
         edu0 = educ_orig_f) %>% 
  filter(!is.na(t), !is.na(w)) %>% 
  mutate(ola = factor(ola, levels = c("1", "2", "3"),
                      labels = c("2016", "2018", "2023")))

mp_hl <- 
  lm_robust(y ~ t + edad + sexo + nac + etnia + hogar + edu0 + ola,
            weights = w,
            se_type = "CR2",
            clusters = idencuesta,
            data = data_hl)

@@

conflicts_prefer(dplyr::filter)

df_pension <- bind_rows(
  tidy(mp_lm, conf.int = TRUE) %>%
    dplyr::filter(term == "t") %>%
    dplyr::mutate(term = paste0("Low-Middle\n", "(N obs. = ", mp_hl$nobs, ";", "\nN clusters = ", mp_hl$nclusters, ")")
) 
  ,
  tidy(mp_lh, conf.int = TRUE) %>%
    dplyr::filter(term == "t") %>%
  dplyr::mutate(term = paste0("Low-High\n", "(N obs. = ", mp_lh$nobs, ";", "\nN clusters = ", mp_lh$nclusters, ")"))
  ,
tidy(mp_mh, conf.int = TRUE) %>%
    dplyr::filter(term == "t") %>%
    dplyr::mutate(term = paste0("Middle-High\n", "(N obs. = ", mp_mh$nobs, ";", "\nN clusters = ", mp_mh$nclusters, ")"))
,
  tidy(mp_ml, conf.int = TRUE) %>%
    dplyr::filter(term == "t") %>%
  dplyr::mutate(term = paste0("Middle-Low\n", "(N obs. = ", mp_ml$nobs, ";", "\nN clusters = ", mp_ml$nclusters, ")"))
,
  tidy(mp_hm, conf.int = TRUE) %>%
    dplyr::filter(term == "t") %>%
  dplyr::mutate(term = paste0("High-Middle\n", "(N obs. = ", mp_hm$nobs, ";", "\nN clusters = ", mp_hm$nclusters, ")"))
,
  tidy(mp_hl, conf.int = TRUE) %>%
    dplyr::filter(term == "t") %>%
  dplyr::mutate(term = paste0("High-Low\n", "(N obs. = ", mp_hl$nobs, ";", "\nN clusters = ", mp_hl$nclusters, ")"))
)

df_pension_lab <- df_pension %>%
  dplyr::mutate(
    sig     = ifelse(conf.low * conf.high > 0, "*", ""),  # IC no cruza 0
    est_lab = sprintf("%.2f%s", estimate, sig),
    low_lab = sprintf("%.2f", conf.low),
    high_lab= sprintf("%.2f", conf.high)
  )

df_pension_lab <- df_pension_lab %>%
  dplyr::mutate(
    term = factor(
      term,
      levels = term
      )
  )

ggplot(df_pension_lab, aes(x = estimate, y = term)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "darkred", size = 0.5, alpha = 1) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0, size = 0.7) +
  geom_point(size = 2.5) +
  scale_x_continuous(limits = c(-0.5, 0.5), breaks = seq(-0.5, 0.5, 0.25)) +
  #geom_text(aes(label = est_lab), hjust = -0.05, size = 4, vjust = -0.5) +
  labs(x = "Estimate", y = NULL) +
  theme_ggdist() 
```

Results reveal a clear directional asymmetry between upward and downward mobility trajectories. Among upwardly mobile respondents, only the Middle→High trajectory exhibits a significant positive effect on support for pension commodification ($\beta$ = 0.09, 95% CI [0.02, 0.16]). Interpreted as an origin-specific ATT, this estimate compares—for individuals who actually moved from the middle to the high class—their observed endorsement with the counterfactual endorsement they would have expressed had they remained immobile in the middle class. Holding all other predictors constant, the point estimate implies a 9 percentage point higher probability of endorsing market-based access to pension benefits for Middle→High movers relative to their own immobility counterfactual, statistically significant at the 95% confidence level. In contrast, two downward trajectories: High→Middle ($\beta$ = −0.07, CI [−0.15, −0.00]) and High→Low ($\beta$ = −0.10, CI [−0.18, −0.02]), show significant negative effects, suggesting that individuals who experience downward mobility from higher-class origins express weaker preferences for market provision. The remaining pathways (Low→Middle, Low→High, Middle→Low) display no significant differences relative to their non-mobile counterparts. These patterns indicate that upward movement within the upper segment (Middle→High) reinforces pro-commodification attitudes, whereas downward movement from privileged origins reduces them.

@@

To probe potential mechanisms underlying the mobility effects, I estimate origin-specific ATT models that interact the mobility treatment with high meritocratic perceptions (ref. = low), using stabilized IPW–ATT weights, covariate adjustment, and wave fixed effects. I also estimate an analogous specification interacting mobility with survey waves (ref. = 2016) to assess temporal heterogeneity in the treatment effect (see [Supplementary Material](#anexo) for complete models).

```{r}
#| echo: false
#| results: false

# 3.4 Interacciones -------------------------------------------------------

# -----------
# Meritocracy 
# -----------

### Interacciones

### Low to middle

data_lm_i <- datos1 %>% 
  select(idencuesta, 
         ola, 
         y = just_pension_f, 
         t = moblm, 
         w = w_low_lm_t, 
         edad = m0_edad_f, 
         sexo = sexo_f, 
         nac = nacionalidad_f,
         etnia = etnia_f,
         hogar = hogar_bip_f, 
         edu0 = educ_orig_f,
         m = merit_i) %>% 
  filter(!is.na(t), !is.na(w)) %>% 
  mutate(ola = factor(ola, levels = c("1", "2", "3"),
                      labels = c("2016", "2018", "2023")))

data_lm_i %>% 
  arrange(idencuesta)

mp_lm_i <-
  lm_robust(y ~ t*m + edad + sexo + nac + etnia + hogar + edu0 + ola,
            weights = w,
            se_type = "CR2",
            clusters = idencuesta,
            data = data_lm_i)

mp_lm_i %>%  
  screenreg()

### Low to high

data_lh_i <- datos1 %>% 
  select(idencuesta, 
         ola, 
         y = just_pension_f, 
         t = moblh, 
         w = w_low_lh_t, 
         edad = m0_edad_f, 
         sexo = sexo_f, 
         nac = nacionalidad_f,
         etnia = etnia_f,
         hogar = hogar_bip_f, 
         edu0 = educ_orig_f,
         m = merit_i) %>%  
  filter(!is.na(t), !is.na(w)) %>% 
  mutate(ola = factor(ola, levels = c("1", "2", "3"),
                      labels = c("2016", "2018", "2023")))

mp_lh_i <- 
  lm_robust(y ~ t*m + edad + sexo + nac + etnia + hogar + edu0+ ola,
            weights = w,
            se_type = "CR2",
            clusters = idencuesta,
            data = data_lh_i)

mp_lh_i %>% 
  screenreg()

### middle to high

data_mh_i <- datos2 %>% 
  select(idencuesta, 
         ola, 
         y = just_pension_f, 
         t = mobmh, 
         w = w_mid_mh_t, 
         edad = m0_edad_f, 
         sexo = sexo_f, 
         nac = nacionalidad_f,
         etnia = etnia_f,
         hogar = hogar_bip_f, 
         edu0 = educ_orig_f,
         m = merit_i) %>% 
  filter(!is.na(t), !is.na(w)) %>% 
  mutate(ola = factor(ola, levels = c("1", "2", "3"),
                      labels = c("2016", "2018", "2023")))

mp_mh_i <-  
  lm_robust(y ~ t*m + edad + sexo + nac + etnia + hogar + edu0+ ola,
            weights = w,
            se_type = "CR2",
            clusters = idencuesta,
            data = data_mh_i)

mp_mh_i %>% 
  screenreg()

### middle to low

data_ml_i <- datos2 %>% 
  select(idencuesta, 
         ola, 
         y = just_pension_f, 
         t = mobml, 
         w = w_mid_ml_t, 
         edad = m0_edad_f, 
         sexo = sexo_f, 
         nac = nacionalidad_f,
         etnia = etnia_f,
         hogar = hogar_bip_f, 
         edu0 = educ_orig_f,
         m = merit_i) %>% 
  filter(!is.na(t), !is.na(w)) %>% 
  mutate(ola = factor(ola, levels = c("1", "2", "3"),
                      labels = c("2016", "2018", "2023")))

mp_ml_i <- 
  lm_robust(y ~ t*m + edad + sexo + nac + etnia + hogar + edu0+ ola,
            weights = w,
            se_type = "CR2",
            clusters = idencuesta,
            data = data_ml_i)

mp_ml_i %>% 
  screenreg()

### High to middle

data_hm_i <- datos3 %>% 
  select(idencuesta, 
         ola, 
         y = just_pension_f, 
         t = mobhm, 
         w = w_hig_hm_t, 
         edad = m0_edad_f, 
         sexo = sexo_f, 
         nac = nacionalidad_f,
         etnia = etnia_f,
         hogar = hogar_bip_f, 
         edu0 = educ_orig_f,
         m = merit_i) %>% 
  filter(!is.na(t), !is.na(w)) %>% 
  mutate(ola = factor(ola, levels = c("1", "2", "3"),
                      labels = c("2016", "2018", "2023")))

mp_hm_i <- 
  lm_robust(y ~ t*m + edad + sexo + nac + etnia + hogar + edu0+ ola,
            weights = w,
            se_type = "CR2",
            clusters = idencuesta,
            data = data_hm_i)

mp_hm_i %>% 
  screenreg()

### High to low

data_hl_i <- datos3 %>% 
  select(idencuesta, 
         ola, 
         y = just_pension_f, 
         t = mobhl, 
         w = w_hig_hl_t, 
         edad = m0_edad_f, 
         sexo = sexo_f, 
         nac = nacionalidad_f,
         etnia = etnia_f,
         hogar = hogar_bip_f, 
         edu0 = educ_orig_f,
         m = merit_i) %>% 
  filter(!is.na(t), !is.na(w)) %>% 
  mutate(ola = factor(ola, levels = c("1", "2", "3"),
                      labels = c("2016", "2018", "2023")))

mp_hl_i <- 
  lm_robust(y ~ t*m + edad + sexo + nac + etnia + hogar + edu0 + ola,
            weights = w,
            se_type = "CR2",
            clusters = idencuesta,
            data = data_hl_i)

mp_hl_i %>% 
  screenreg()
```

@@

df_pension_i <-  bind_rows(
  tidy(mp_lm_i, conf.int = TRUE) %>%
    filter(term %in% c("t:mHigh", "t")) %>%
    mutate(name = paste0("Low-Middle\n", "(N obs. = ", mp_hl_i$nobs, ";", "\nN clusters = ", mp_hl_i$nclusters, ")")
    ) 
  ,
  tidy(mp_lh_i, conf.int = TRUE) %>%
    filter(term %in% c("t:mHigh", "t")) %>%
    mutate(name = paste0("Low-High\n", "(N obs. = ", mp_lh_i$nobs, ";", "\nN clusters = ", mp_lh_i$nclusters, ")"))
  ,
  tidy(mp_mh_i, conf.int = TRUE) %>%
    filter(term %in% c("t:mHigh", "t")) %>%
    mutate(name = paste0("Middle-High\n", "(N obs. = ", mp_mh_i$nobs, ";", "\nN clusters = ", mp_mh_i$nclusters, ")"))
  ,
  tidy(mp_ml_i, conf.int = TRUE) %>%
    filter(term %in% c("t:mHigh", "t")) %>%
    mutate(name = paste0("Middle-Low\n", "(N obs. = ", mp_ml_i$nobs, ";", "\nN clusters = ", mp_ml_i$nclusters, ")"))
  ,
  tidy(mp_hm_i, conf.int = TRUE) %>%
    filter(term %in% c("t:mHigh", "t")) %>%
    mutate(name = paste0("High-Middle\n", "(N obs. = ", mp_hm_i$nobs, ";", "\nN clusters = ", mp_hm_i$nclusters, ")"))
  ,
  tidy(mp_hl_i, conf.int = TRUE) %>%
    filter(term %in% c("t:mHigh", "t")) %>%
    mutate(name = paste0("High-Low\n", "(N obs. = ", mp_hl_i$nobs, ";", "\nN clusters = ", mp_hl_i$nclusters, ")"))
)

df_pension_lab_i <- df_pension_i %>%
  mutate(
    sig     = ifelse(conf.low * conf.high > 0, "*", ""),  # IC no cruza 0
    est_lab = sprintf("%.2f%s", estimate, sig),
    low_lab = sprintf("%.2f", conf.low),
    high_lab= sprintf("%.2f", conf.high)
  )

df_pension_lab_i <- df_pension_lab_i %>%
  mutate(
    name = factor(
      name,
      levels = unique(name)
    ),
    term = if_else(term == "t", "Low Meritocracy", "High Meritocracy"),
    term = factor(term, levels = c("High Meritocracy", "Low Meritocracy"))
  )

ggplot(df_pension_lab_i, aes(x = estimate, y = name)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "darkred", size = 0.5, alpha = 1) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0, size = 0.7) +
  geom_point(size = 2.5) +
  scale_x_continuous(limits = c(-0.5, 0.5), breaks = seq(-0.5, 0.5, 0.25)) +
  facet_wrap(~term) + 
  #geom_text(aes(label = est_lab), hjust = -0.05, size = 4, vjust = -0.5) +
  labs(x = "Estimate", y = NULL) +
  theme_ggdist()

@@

# -----------
# Time 
# -----------

# Low to middle

mp_lm_t <- lm_robust(y ~ t*ola + edad + sexo + nac + etnia + hogar + edu0,
          weights = w,
          se_type = "CR2",
          clusters = idencuesta,
          data = data_lm)

# Low to high

mp_lh_t <- lm_robust(y ~ t*ola+ edad + sexo + nac + etnia + hogar + edu0,
          weights = w,
          se_type = "CR2",
          clusters = idencuesta,
          data = data_lh)

# Middle to high

mp_mh_t <- lm_robust(y ~ t*ola + edad + sexo + nac + etnia + hogar + edu0,
          weights = w,
          se_type = "CR2",
          clusters = idencuesta,
          data = data_mh)

# Middle to low

mp_ml_t <- lm_robust(y ~ t*ola + edad + sexo + nac + etnia + hogar + edu0,
          weights = w,
          se_type = "CR2",
          clusters = idencuesta,
          data = data_ml)

# High to middle

# tiempo: descender se vuelve un efecto positivo en la ultima ola (con mi plata no)
mp_hm_t <- lm_robust(y ~ t*ola + edad + sexo + nac + etnia + hogar + edu0,
          weights = w,
          se_type = "CR2",
          clusters = idencuesta,
          data = data_hm)

# High to low

mp_hl_t <- lm_robust(y ~ t*ola + edad + sexo + nac + etnia + hogar + edu0,
          weights = w,
          se_type = "CR2",
          clusters = idencuesta,
          data = data_hl)

>>>>

@@

```{r}
#| label: source1
#| echo: false
#| results: false
#| message: false
#| warning: false
#| include: false

library(here)
source(here("processing/prod_analysis_multinom.R"))
```

@@

```{r}
#| label: fig-alluvial
#| fig-cap: "Change in preferences for pension commodification over time (2016, 2018, and 2023)."
#| fig-cap-location: bottom
#| echo: false
#| results: asis
g1
```

@@

g2

@@

@fig-reg reports the doubly robust estimates of the effect of intergenerational mobility on preferences for pension commodification, obtained from weighted models combining stabilized IPW with covariate adjustment and wave fixed effects. Results reveal a clear directional asymmetry between upward and downward mobility trajectories. Among upwardly mobile respondents, only the Middle→High trajectory exhibits a significant positive effect on support for pension commodification ($\beta$ = 0.09, 95% CI [0.02, 0.16]). Interpreted as an origin-specific ATT, this estimate compares—for individuals who actually moved from the middle to the high class—their observed endorsement with the counterfactual endorsement they would have expressed had they remained immobile in the middle class. Holding all other predictors constant, the point estimate implies a 9 percentage point higher probability of endorsing market-based access to pension benefits for Middle→High movers relative to their own immobility counterfactual, statistically significant at the 95% confidence level. In contrast, two downward trajectories: High→Middle ($\beta$ = −0.07, CI [−0.15, −0.00]) and High→Low ($\beta$ = −0.10, CI [−0.18, −0.02]), show significant negative effects, suggesting that individuals who experience downward mobility from higher-class origins express weaker preferences for market provision. The remaining pathways (Low→Middle, Low→High, Middle→Low) display no significant differences relative to their non-mobile counterparts. These patterns indicate that upward movement within the upper segment (Middle→High) reinforces pro-commodification attitudes, whereas downward movement from privileged origins reduces them.

@@

g3
```

@@

To probe potential mechanisms underlying the mobility effects, I estimate origin-specific ATT models that interact the mobility treatment with high meritocratic perceptions (ref. = low), using stabilized IPW weights, covariate adjustment, and wave fixed effects. I also estimate an analogous specification interacting mobility with survey waves (ref. = 2016) to assess temporal heterogeneity in the treatment effect (see [Supplementary Material](#anexo) for complete models).

@@

g4
```

@@
